---
title: "Static App Prototype"
author: "Andy Powers"
format: pdf
---

# Introduction

## Static App Prototype
This Quarto R project is the static, phase-I build of my app for exploring mobile device data. After completion in static form, I will iteratively convert each element to R Shiny dynamic format and add a flexible and friendly interface.

## Static App Components
1. Mobile device dataset.
2. Contingency tables.
3. Numerical summaries.
4. Plots.

# Components

## Libraries
```{r}
library(tidyverse)
library(utils)
library(readr)
library(readxl)
```

## Dataset

This app will explore mobile device usage data from [Kaggle](https://www.kaggle.com/datasets/valakhorasani/mobile-device-usage-and-user-behavior-dataset). Because Kaggle is difficult to query, I've saved the file locally and included it in the same directory as the Quarto and app files.

```{r}
data <- readr::read_csv("user_behavior_dataset.csv",col_names = TRUE)
```

## Contingency Tables
A starting place for analysis, let's look at some one- and two-way contingency tables.
```{r}
data |>
  group_by(`Device Model`) |>
  summarize(n())
```

So, our dataset is a fairly even mix of samples across device models.

```{r}

```

